---
title: "GSS challange"
author: "Kimberly Lemus Munoz"
date: "Summer 2022"
output: html_document
---







```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2018.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,sex,race,cappun,degree,relig,gunlaw,tvhours,owngun,age,wordsum)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->



## Now with LESS data (more variables give less data due to NA's)

```{r}
dd2 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The first graph above, which includes the counts of people who are in favor of and are opposed to capitol punishment, shows that about 350 out of the about 500 white people are in favor of capitol punishment. The second graph shows that  more than 65% of the group are in favor of capital punishment. On the other hand, more than half of black people are opposed to capitol punishment while more than half of other racial groups are in favor of it.  

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->




## Now with less data

```{r}
table2 <- xtabs(~race + cappun, data=mydata)
rowPerc(table2)
colPerc(table2)
```
The first table compares the percentages of people who are in favor of and are opposed to capital punishment per racial group. It specifies that 67.98% of white people are in favor of capital punishment while 32.02% are opposed to it. Meanwhile, 46.29% of black people are in favor of capitol punishment while 53.71% are opposed to it. For other racial groups, however, 55.87% are in favor while 44.13% are opposed. Therefore, the first table shows that white and other racial groups are generally in favor of capital punishment while black people are opposed to it.
  
The second table, on the other hand, shows the percentages of racial groups that make up the population of people who are in favor of and are opposed to capitol punishment, respectively. Because there are more white people compared to other racial groups, they make up most of the population who are in favor and are opposed to capitol punishment, with 78.34% and 63.24%, respectively. And because there are less number of black people and others in the sample, they make up less of the population who are in favor and are opposed to capitol punishment.    
  

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

`


## Less data version

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
#fisherXY(mydata$race,mydata$cappun)
```
```{r}
fisherXY = function(x, y, confL = 0, printAll=TRUE){
  # Inputs: 
  #   x: factor variable, 2 levels
  #   y: factor variable, n levels
  # confL: confidence level, default = 0.95
  #        print only those tables, results if p-value < (1-confL)
  # printAll: overide the 'print only if p-value < (1-confL), and
  #           print all tables and fisher test results
  #
  # Constraint:
  #   x and y must be same length
  #
  # Outputs:
  #   All 2x2 Fisher exact tests
  
  if(confL<0 || confL> 1){
    print('error: confL must be between 0 and 1')
    return(999)
  }
  
  if(!is.factor(x)) x=as.factor(x)
  if(!is.factor(y)) y=as.factor(y)
  
  fisherTable = as.matrix(table(x, y))
  cat('\ncontingency table:\n')
  print(fisherTable)
  cat('\n***All 2X2 tests****\n\n')

  # all 2x2 tests
  levelsY = levels(y)
  levelsX = levels(x)
  
  m = length(levelsX)
  n = length(levelsY)
  
  confLevel = 0.95
  if (confL !=0 ) confLevel=confL
  
  for (i in 1:(m-1))
       for(j in ((i+1):m))
          for (k in 1:(n-1))
             for (l in ((k+1):n)) {
                
                 fisherIndexX = c(i,j)
                 fisherIndexY = c(k,l)
                 
                 fTempTable = as.matrix(fisherTable[fisherIndexX,
                                        fisherIndexY])
                 
                 fisherTest=fisher.test(fTempTable, conf.int = TRUE, 
                              conf.level = confLevel)
                 
                 if((fisherTest$p.value < (1-confLevel) ) || printAll){
                   
                       print(c(levelsX[i], levelsX[j]))
                       print(c(levelsY[k], levelsY[l]))
                       print(fTempTable)
                       print(fisherTest)
                       cat('\n')
                 }
                 
        }
}

```

Based on the Chi’s squared test above, the p-value is low at 1.027e-14. This means that we can reject the null hypothesis that states that attitude towards capital punishment is independent of race. To further elaborate on the data and compare each racial groups’ attitude, a series of fisher’s exact tests are conducted. This p-value is comparably larger than than the one from the master branch, and this is because this branch has less data which will therefore show less difference. 


```{r}
fisherXY(mydata$cappun,mydata$race)
```
 
 Based on the Fisher XY tests above, which compare two racial groups at a time, we can tell that the attitudes of white and black people toward capital punishment are different based on the low p-value of 5.75e-14. The attitudes of white and others toward capital punishment, on the other hand, are the same based on the p-value of 1. And lastly, the attitudes of black people are others are also different based on the low p-value of 0.02479.
  
  
Therefore, some of the racial groups' attitudes towards capital punishment are different from the rest. However, comparing this with the branch with more data, there is less difference in terms of attitude towards capital punishment due to the smaller number of data in this branch. Nevertheless, the data above still support the claim that attitude towards capital punishment is dependent on race. 
 
 
 
 
 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->



The graphical and numerical descriptive results, along with the inferential results, all show that attitude towards capital punishment is dependent on race. White, black, and other racial groups provided varying responses on whether they are in favor of or are opposed to capital punishment. Therefore, we reject the null hypothesis and claim that there is a relationship between attitude towards capital punishment and race.

Comparing it to the other data set of the other branch, this difference in attitude towards capital punishment is smaller due to the fact that this data set contains less data points than the other. Generally, less data points would provide less difference, and this is exhibited by this data set.

